---
title: "Function_Examples"
author: "Andrea Lanz"
date: "12/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this document is to provide examples for generating music using the functions I have created.

### Simple random notes:
The first example will show how to generate music using the simple random sample method:

First generate a sample:
```{r}
n <- 5 #assign the number of measures to output
sample1 <- generate_simple_sample(n)
```

Create the file, which will be placed in the specified path:
```{r}
generate_lilypond(sample1, n, file_name = "results/example1.ly")
```

And the function with plot the notes, which can be diabled with an argument.

### Simple random notes with rhythm:
The second example will show how to generate a song with more variation in speed.

Generate the sample, specifying the number of measures:
```{r}
m <- 5 #assign the number of measures to output
sample2 <- generate_rhythm_sample(m)
```

Create the file, which will be placed in the specified path:
```{r}
generate_lilypond(sample2, m, file_name = "results/example2.ly")
```

The notes are as shown above.

### Rhythm and notes with consonance:
